---
title: "Creating annotatable plots: the 'annoplot' object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{annoplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shiny)
library(annoplots)
library(ggplot2)
library(dplyr)
library(purrr)

```

# A minimal reproducible example

This R package helps to create annotatable plots in rShiny. For example, after selecting points in a plot, you may wish to change them to a highlighted color, label them with relevant information, or zoom in on those points. This is done by the creation of a new [Shiny module](https://shiny.rstudio.com/articles/modules.html), `[MODULE NAME]`, which displays the plots and reads the points selected by the user, and an `AnnoPlot` object, which contains all plot-related data and annotates the plots. 


Here I use the `mtcars` data set to create a single base plot:

```{r}


p_base <- ggplot(mtcars, aes(x = mpg, y = cyl)) + geom_point()
p_base
```




When I select points from the base plot in the Shiny app, I want those points to turn red. I specify this using the annotation function which must take the arguments `plots` and `pointdf`. By passing the selected points as a dataframe in `pointdf`, I can specify which points to apply the annotation to.

```{r}
library(purrr)
af_highlightred <- function(plots, pointdf) {
    map(plots, ~. + geom_point(data = pointdf, color = 'red') )
}

# Take a random subset of points
selected_points <- sample(1:nrow(mtcars), 3)
pointdf <- mtcars[selected_points, ]

af_highlightred(list(p_base), pointdf)

```


Or maybe I want to color the points red and label with the model of the car.

```{r}
library(ggrepel)

mtcars_labeled <- mutate(mtcars, model = rownames(mtcars))
pointdf <- mtcars_labeled[selected_points, ]


af_highlightlabel <- function(plots, pointdf) {
    map(plots, ~. + geom_point(data = pointdf, color = 'red') +
          geom_text_repel(aes(label = model), data = pointdf))
}

af_highlightlabel(list(p_base), pointdf)
```

To create an object that the Shiny modules will recognize and be able to communicate with, I create an  `AnnoPlot` object which contains the base plot, annotation function, data, and other configurable options (see [annoplot link]).

```{r}
ap_mtcars <- annoplot(plots = list(p_base), 
                      plotdata = mtcars_labeled, 
                      annofun = af_highlightlabel) 
# returns an `AnnoPlot` object
```


This is essentially a standardized wrapper around the actions we just did: displaying the base plot and carrying out the annotation function. 

```{r}
ap_plots(ap_mtcars) # returns the list of base plots
ap_highlight(ap_mtcars, pointdf) # returns annotated plots
ap_get(ap_mtcars, "annotype") # returns the 'annotype' option 
ap_set(ap_mtcars, "annotype", "click") # user selects points by clicking them
```


This is then relayed to the AP [Shiny module](https://shiny.rstudio.com/articles/modules.html), which is called using two functions: 

- `APOutput()`

- `APServer()`


```{r}
# ui <- fluidPage(
# 
#   # App title ----
#   titlePanel("A minimal AnnoPlot example"),
#     # Main panel for displaying outputs ----
#     mainPanel(
# 
#       # Output: Histogram ----
#       plotOutput(id = "mtcars")
# 
#     )
# )
# 
# 
# server <- function(input, output) {
#   points <- reactiveValues(pt_idx = c())
#   selected <- APServer("mtcars", ap_mtcars,
#                        reactive(c(input$button_go, input$button_clear)),
#                        points, scale = 4, hover = T)
#   
#   
#    observeEvent(selected(), {
#         observe("points from top plots")
#         print(top())
#         points$pt_idx <- unique(c(points$pt_idx, top()))
#     })
# 
# }
# 
# shinyApp(ui, server)
```




# Updating multiple plots 



# Communicating across multiple AP Shiny modules



# Modifying the Shiny module for your own plots



