---
title: "Example: Radon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: Radon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(rstanarm)
library(ssbf)
library(annoplots)
library(tidyverse)
theme_set(theme_bw(base_size = 13))
```


NEED TO MOVE THIS OUT OF THIS PACKAGE. But for now load slice, then annoplots

```{r}
contrasts_for_pooling <- function(df) {
    categorical_mask <- purrr::map_lgl(df, ~is.character(.) | is.factor(.))
    categorical_names <- colnames(df)[categorical_mask]
    purrr::map(categorical_names, ~stats::contrasts(df[[.]], contrasts=F)) %>%
        purrr::set_names(categorical_names)
}
radon_v <- function(rstfit, X) {
    samps <- rstan::extract(rstfit$stanfit)
    sigs <- list("sigma2" = mean(samps$theta_L), "tau2" = mean(samps$aux)^2)
    
    cty_mask <- str_detect(colnames(X), "county")
    sigidiag <- rep(0, ncol((X)))
    sigidiag[cty_mask] <- 1/sigs$sigma2
    
    siginv <- diag(sigidiag)
    
    solve(t(X) %*% X/sigs$tau2 + siginv) %*% t(X)/sigs$tau2
    
    
}


radon_w <- function(rstfit, X) {
    
    X %*% radon_v(rstfit, X)  
    
}
radon_add_relp <- function(df) {
    df %>%
        group_by(id__, q_county, q_floor) %>%
        mutate(
            relationship = case_when(
                same_county & same_floor ~ "same cluster",
                same_county ~ "same county",
                same_floor ~ "same basemen",
                T ~ "different county and basement"
            ),
            sign = sign(val),
            q_n = length(unique(q_id))
        ) %>% ungroup()
}


radon_add_summaries <- function(relpdf) {
    relpdf %>%
        group_by(id__) %>% 
        mutate(
            ssbf = sum(val[relationship != "same cluster"]^2),
            weight_ratio = val/mean(val[relationship == "same cluster"]),
            pf = sum(val[relationship != "same cluster"])
            
        ) %>%
        ungroup()
    
}

radon_total_bsize <- function(df) {
    df  %>%
        group_by(p_county, p_floor) %>%
        mutate(p_n = length(unique(id__))) %>% 
        ungroup()
}




radon_weight_ptbypt <- function(slice) {
    slice$factor_df() %>% 
        radon_add_relp() %>%
        radon_add_summaries() %>%
        radon_total_bsize() %>%
        dplyr::select(-rowname, -colname, -id__, 
                      -ends_with("_y"), -starts_with("same_"),
                      -starts_with("similar_")) %>%
        unique()
}

radon_label_basement <- function(df) {
    # must have p_floor
    df %>% 
        mutate(
            basement = case_when(p_floor == 0 ~ "B) No basement",
                                 p_floor == 1 ~ "A) Has basement")
        ) 
}

radon_label_sign <- function(df) {
    df %>% mutate(
        sign = case_when(sign < 0 ~ "neg", T ~ "pos") %>%
            factor(ordered = T, levels = c("pos", "neg"))
    )
}

summarise_weights <- function(grpeddf) {
    grpeddf %>% summarise(
        mean_val = mean(val), 
        q_n = length(unique(q_id)), 
        ratio = mean(weight_ratio),
        total = sum(val), 
        unique_signs = length(unique(sign(val))),
        n_pos = sum(val >= 0),
        n_neg = sum(val < 0),
        pssbf = sum(val^2)
    ) 
}

radon_summarise_ptbypt <- function(ptbypt) {
    ptbypt %>%
        radon_label_basement() %>%
        radon_label_sign() %>%
        group_by_at(vars(starts_with("p_"), pf, relationship, sign, basement, ssbf)) %>%
        summarise_weights()
}

```

# Data

Description of data and model



```{r mod3wts}

set.seed(29873)
radonfit3 <- rstanarm::stan_glmer(
    log_radon ~ floor + log_uranium + (1|county), 
    data = rstanarm::radon, refresh = 0
)
radon_dat <- mutate(rstanarm::radon, county = factor(county))
X3 <- model.matrix(~floor + log_uranium + county, 
                   data = radon_dat,
                   contrasts = contrasts_for_pooling(radon_dat))

W <- radon_w(radonfit3, X3[,-3])
slice_mod3_post <- slice::BasicSlice$new(
    dplyr::select(radon_dat, -log_radon) %>% mutate(floor = factor(floor)),
    W, y = as.vector(radon_dat$log_radon), 
    yhat = as.vector(W %*% radon_dat$log_radon)
)


```


# Generate plots

## Define relationship groups
```{r}
ptbypt_3 <- radon_weight_ptbypt(slice_mod3_post)
grp_labels <- expression(b[kj],  b[k*minute~j], b[kj*minute])
grp_colors <- c("black", RColorBrewer::brewer.pal(3, "Dark2"))
```


## Determine quantities to plot


```{r}
indvwt_3 <- radon_summarise_ptbypt(ptbypt_3)

pdf <-  ungroup(indvwt_3) %>% select(
    p_county, p_log_uranium, basement, 
    relationship, ssbf, total, pf, p_n, q_n, ratio
) %>%
    filter(str_detect(relationship, "same")) %>%
    mutate(
        id__ = as.numeric(factor(interaction(p_county, basement))),
        relationship = factor(relationship, 
                              levels = c("same cluster", "same basemen", "same county"),
                              labels = grp_labels)
    )
```


## Define base plots and annotation functions

```{r, fig.height = 12, fig.width = 5}

p_data <- ggplot(pdf, aes(x = basement, y = p_county, fill = p_n)) +
    geom_tile() +
    labs(title = expression(n[kj]), x = NULL, y = NULL, fill = NULL) +
    scale_x_discrete(position = 'top') +
    scale_fill_distiller(palette = "Blues", direction = 1, trans = 'log', 
                         breaks = c(1, 2, 4, 7, 25, 100), labels = round) +
    theme(legend.position = "top") +
    scale_y_discrete(limits = rev)
p_data
af_addpoint <- function(plots, pointdf) {
    purrr::map(plots, ~. + geom_point(data = pointdf) )
}


ap_pn <- annoplot(
    # required
    p_data, 
    select(pdf, basement, p_county, p_n, ssbf, id__) %>% unique(),
    af_addpoint,
    # extra
    hover = c("ssbf"), dims = list(c(6, 1.75)
    ))
```


```{r}
p_lendersize <- ggplot(pdf, aes(x = q_n, y = ssbf)) + 
    geom_point() +
    facet_grid(~relationship, labeller = "label_parsed") +
    scale_x_continuous(trans = 'log', labels = round) +
    labs(x = "Relationship group size", y = "SSBF", title = "Cluster size")

af_highlightred <- function(plots, pointdf) {
    purrr::map(plots, ~. + geom_point(data = pointdf, color = 'red') )
}


ap_qn <- annoplot(
    p_lendersize,  
    select(pdf, p_county, p_log_uranium, basement, relationship, ssbf, q_n, id__) %>% unique(),
    af_highlightred,
    hover = c("p_county", "basement", "p_log_uranium"), dims = list(c(1, 4)
                                                                    
    ))
```




```{r}
p_relps_base <- ggplot(pdf, aes(x = total, y = ssbf, color = relationship, shape = relationship)) +
    facet_grid(~basement) +
    scale_color_manual(
        values = grp_colors,
        labels = grp_labels
    ) +
    scale_shape_manual(values = c(19, 17, 15), labels = grp_labels) +
    labs(x = "Borrowing factor", y = "SSBF", color = NULL, shape = NULL,
         title = "BFs") +
    theme(legend.position = 'top')



ap_relps <- annoplot_fade(p_relps_base, geom_point, pdf, plot_dims = list(c(1, 2)),
                                hover = c("p_county", "ssbf", "total"))
```



```{r}
p_relps_ratio <- ggplot(pdf, aes(x = ratio, y = ssbf, color = relationship, shape = relationship)) +
    facet_grid(~basement) +
    scale_color_manual(
        values = grp_colors,
        labels = grp_labels
    ) +
    scale_shape_manual(values = c(19, 17, 15), labels = grp_labels) +
    theme_bw() +
    labs(x = "BF/SF", y = "SSBF", color = NULL, shape = NULL, title = "Ratio of BF to SF") +
    theme(legend.position = 'none')

ap_relps_ratio <- annoplot_fade(p_relps_ratio, geom_point, pdf, plot_dims = list(c(1, 2)),
                                hover = c("p_county", "ssbf", "ratio"))
```



# Run Shiny app



```{r, eval = F}
sso <- SLICE_app$new(pdf %>% select(p_county, basement, id__, ssbf, p_n, pf) %>% unique(),
                     title = "SSBF: Radon",
                     factorplots = list(ap_relps, ap_relps_ratio),
                     sideplots = list(ap_pn),
                     errorplots = list(ap_qn))
ssbf::launch(sso)
```

